\normalfont\spacedallcaps{Hazardous asteroids forecast via Markov random fields}

### 2021-12-03 Manuscript/manuscript.tex
PUB

Let us now move on to the mixed interaction model, where the hazardousness feature was also considered.  The first model was calculated with the mgm package  \cite{mgm,haslbeck2015mgm} where the nodewise regression is used \cite{meinshausen2006high}. The $k$ parameter was set equal to two, and a cross-validation (CV) with ten folds was considered. The result is reported in Fig. \ref{mgm}. From this figure, it can be seen that the features that are directly connected to the Hazardous feature are: the absolute magnitude, the min orbit intersection, and the eccentricity.  The theory supports all these connections. The min orbit intersection is the main parameter for the hazardous value. On the other side, the eccentricity can be tough as a parameter that describes how near the celestial body moves at the perihelion (indeed, this parameter is correctly linked with the perihelion distance). The Absolute Magnitude is also a meaningful parameter for the hazard evaluation since, if the asteroid is too small, it will be destroyed by the earth atmosphere.  Furthermore, the parameters related to the diameters are also not connected with the orbital parameters. In addition, the relative velocity is not directly related to the mean motion, but there are the orbital parameters within as expected. Thus it can be said that the model produced is almost consistent with the astronomic laws. On this basis, the performances of the model in terms of confusion matrix, ROC (Receiver operating characteristic) curve and $\phi$ coefficient (also known as Matthews correlation coefficient ) were evaluated (for this purpose the Caret R package \cite{kuhn2008building,caret} was used). The confusion matrix obtained is reported in Fig. \ref{mgm_confusion}: the ROC curve is given in Fig. \ref{ROC_mgm} and it corresponds to a $\phi=0.6$. These performances will be commented on in the next section when the performances of other ML algorithms for the same dataset will be reported. Beside the mgm algorithm/package other approaches were also considered for the evaluation of the mixed interaction graphical model: the function \textit{mmod()} of gRim package \cite{hojsgaard2012graphical} and the \textit{minforest()} of gRapHD (which uses the minForest method) \cite{de2009high}. For both cases, a stepwise algorithm was used.  Their result are reported in Fig. \ref{mmod} and \ref{minforest}. It can be seen that in these models, the minimum diameter is linked with the hazardous feature, which is correct, but what is without a physical meaning is that this quantity is also linked to the features connected with the orbital parameters. A weak link with the orbital parameters is meaningful for the magnitude (as the mgm model for the mean anomaly).  The magnitude depends on the distance between the asteroids and the observer.   However, a direct link from the volume parameters (and thus from the mass) with the orbital ones is not acceptable. In principle, one can set these links as forbidden in the algorithm (blacklist), but in the author view, a model where the result is obtained without boundaries should be preferred to those obtained with a large number of boundaries. Therefore we reject the mmod and minforest ones favouring the mgm, which was obtained without boundaries.

@@

We see that the two orbital parameters $T'$ and $a'$ are independent to orbiting mass. This statement can be extended to the other orbital parameters by considering the previous approximation $m$,$m'<<m_{c}$. This is why we expect that the mass/volume of the asteroid can not be conditionally dependant on the orbital parameters. It is useful to define also the \textcolor{red}{mean motion} (feature that is also present in the asteroids dataset) as \cite{murray1999solar}:

>>>>

Let us now move on to the mixed interaction model, where the hazardousness feature was also considered.  The first model was calculated with the mgm package  \cite{mgm,haslbeck2015mgm} where the nodewise regression is used \cite{meinshausen2006high}. The $k$ parameter was set equal to two, and a cross-validation (CV) with ten folds was considered. The result is reported in Fig. \ref{mgm}. From this figure, it can be seen that the features that are directly connected to the Hazardous feature are: the absolute magnitude, the min orbit intersection, and the eccentricity.  The theory supports all these connections. The min orbit intersection is the main parameter for the hazardous value. On the other side, the eccentricity can be tough as a parameter that describes how near the celestial body moves at the perihelion (indeed, this parameter is correctly linked with the perihelion distance). The Absolute Magnitude is also a meaningful parameter for the hazard evaluation since, if the asteroid is too small, it will be destroyed by the earth atmosphere.  Furthermore, the parameters related to the diameters are also not connected with the orbital parameters. In addition, the relative velocity is not directly related to the mean motion, but there are the orbital parameters within as expected. Moving to the relation of orbital parameters, it can be seen that there is a negative relation between the eccentricity and perihelion distance: this is meaningful since it comes from the definition of eccentricity. The other side of the coin of this connection is the positive relationships within the semimajor axis and the eccentricity. Furthermore, the positive relationship between the mean motion and the semimajor axis is correct: this is the result of the 3$^{th}$ Kepler law. Finally the connections with the Jupiter Tisserand Invariant comes directly from the definition of this parameter. A full explanation of these connections is provided in the Appendix A. Thus it can be said that the model produced is consistent with the astronomic laws. On this basis, the performances of the model in terms of confusion matrix, ROC (Receiver operating characteristic) curve and $\phi$ coefficient (also known as Matthews correlation coefficient ) were evaluated (for this purpose the Caret R package \cite{kuhn2008building,caret} was used). The confusion matrix obtained is reported in Fig. \ref{mgm_confusion}: the ROC curve is given in Fig. \ref{ROC_mgm} and it corresponds to a $\phi=0.6$. These performances will be commented on in the next section when the performances of other ML algorithms for the same dataset will be reported. Beside the mgm algorithm/package other approaches were also considered for the evaluation of the mixed interaction graphical model: the function \textit{mmod()} of gRim package \cite{hojsgaard2012graphical} and the \textit{minforest()} of gRapHD (which uses the minForest method) \cite{de2009high}. For both cases, a stepwise algorithm was used.  Their result are reported in Fig. \ref{mmod} and \ref{minforest}. It can be seen that in these models, the minimum diameter is linked with the hazardous feature, which is correct, but what is without a physical meaning is that this quantity is also linked to the features connected with the orbital parameters. A weak link with the orbital parameters is meaningful for the magnitude (as the mgm model for the mean anomaly).  The magnitude depends on the distance between the asteroids and the observer.   However, a direct link from the volume parameters (and thus from the mass) with the orbital ones is not acceptable. In principle, one can set these links as forbidden in the algorithm (blacklist), but in the author view, a model where the result is obtained without boundaries should be preferred to those obtained with a large number of boundaries. Therefore we reject the mmod and minforest ones favouring the mgm, which was obtained without boundaries.

@@

We see that the two orbital parameters $T'$ and $a'$ are independent to orbiting mass. This statement can be extended to the other orbital parameters by considering the previous approximation $m$,$m'<<m_{c}$. This is why we expect that the mass/volume of the asteroid can not be  dependant on the orbital parameters. It is useful to define also the \textcolor{red}{mean motion} (feature that is also present in the asteroids dataset) as \cite{murray1999solar}:
